Predator clusters

### 2023-03-14 reports/04_pred_clus.qmd
Details for optimal cluster number

We use the gap statistic [@tibshirani2001] to choose the optimal number of clusters ($k$). The gap statistic ($f$) is a goodness of clustering measure. The authors recommended choosing the smallest $k$ such that $f(k) \geq f(k+1) - s_{k+1}$ where $s$ is the standard error of $f$. According to this rule, the predators are best described by two clusters.

```{r}
cluster_fn <- function(x, k) {
  bray_dist <- function(x) {
    vegdist(x, method = "bray")
  }
  list(cluster = cutree(hclust(bray_dist(x), method = "ward.D2"), k = k))
}
sightings_gap <- suppressWarnings(
  cluster::clusGap(sightings_mtx, 
                   FUN = cluster_fn,
                   K.max = 10, 
                   B = 500,
                   verbose = FALSE)
)
sightings_gap_df <- as_tibble(sightings_gap$Tab) %>% 
  mutate(
    k = seq_along(logW),
    is_optimal = k == cluster::maxSE(gap, SE.sim, method = "Tibs2001SEmax")
  )
firstmaxk <- with(sightings_gap_df,
                  cluster::maxSE(gap, SE.sim, method = "firstmax"))
gap_thr <- with(sightings_gap_df, gap[firstmaxk] - SE.sim[firstmaxk])
ggplot(sightings_gap_df, aes(x = k, y = gap)) +
  geom_line() +
  geom_pointrange(aes(ymin = gap - SE.sim, ymax = gap + SE.sim,
                      color = is_optimal)) +
  geom_hline(yintercept = gap_thr, color = "red", linetype = "dashed") +
  scale_color_manual(values = c(`TRUE` = "red",
                                `FALSE` = "black")) +
  theme_classic() +
  theme(legend.position = "none")
```

>>>>

We use the gap statistic [@tibshirani2001] to choose the optimal number of clusters ($k$). The gap statistic ($f$) is a goodness of clustering measure. The authors recommended choosing the smallest $k$ such that $f(k) \geq f(k+1) - s(k+1)$ where $s$ is the standard error of $f$. According to this rule, the predators are best described by two clusters. However, $f(2)$ is only slightly greater than $f(3)-s(3)$ and, because $s$ is estimated by bootstrapping, $k=3$ may be selected sometimes depending on stochasticity.

```{r}
cluster_fn <- function(x, k) {
  bray_dist <- function(x) {
    vegdist(x, method = "bray")
  }
  list(cluster = cutree(hclust(bray_dist(x), method = "ward.D2"), k = k))
}
set.seed(931)
sightings_gap <- suppressWarnings(
  cluster::clusGap(sightings_mtx, 
                   FUN = cluster_fn,
                   K.max = 10, 
                   B = 500,
                   verbose = FALSE)
)
sightings_gap_df <- as_tibble(sightings_gap$Tab) %>% 
  mutate(
    k = seq_along(logW),
    is_optimal = k == cluster::maxSE(gap, SE.sim, method = "Tibs2001SEmax")
  )
ggplot(sightings_gap_df, aes(x = k, y = gap)) +
  geom_line() +
  geom_pointrange(aes(ymin = gap - SE.sim, ymax = gap + SE.sim,
                      color = is_optimal)) +
  scale_color_manual(values = c(`TRUE` = "red",
                                `FALSE` = "black")) +
  theme_classic() +
  theme(legend.position = "none")
```
